---
title: "Exploratory Data Analysis: Hands-on"
author: "Originally written by Bon Woo Koo & Subhro Guhathakurta; modified by Uijeong Hwang"
date: '2024-09-23'
output:
  rmdformats::downcute:
    downcute_theme: "chaos"
    code_folding: show
editor_options: 
  markdown: 
    wrap: 72
---

```{r,message=FALSE, include=FALSE, r,message=FALSE}
tidycensus::census_api_key(Sys.getenv("census_api_key"))
library(tidycensus)
library(sf)
library(tmap)
library(jsonlite)
library(tidyverse)
library(httr)
library(jsonlite)
library(reshape2)
library(here)
library(knitr)
library(skimr)
library(units)
ttm()
```

```{=html}
<style type="text/css">
  body{
  font-family: Arial;
  }
</style>
```
# Yoga Studio and Gentrification

Gentrification is the process of urban transformation in which
neighborhoods undergo socioeconomic changes, typically characterized by
an influx of higher-income residents and investment. This often leads to
increased property values and rents, which can displace lower-income
residents and alter the cultural landscape of the area. While
gentrification can improve infrastructure and amenities, it raises
concerns about social equity, displacement, and the loss of community
identity.

> "Seemingly overnight, a yoga studio replaces a barbershop, a coffee café takes over a small grocery, and a multi-story apartment building looms where older single-family homes once stood." ([Katharine Miller, 2024)](https://hai.stanford.edu/news/spotting-visual-signs-gentrification-scale))

In today's lab session, you will be answering this question: *Are yoga studios associated with gentrification?*

## Operational Definition of Gentrification

Gentrification is about change; the transformation of a neighborhood once affordable becoming unaffordable, so we have two conditions: "*once affordable*" and "*becoming unaffordable*".

To create a simple operational definition, let's say:

* Measure of affordability: real estate tax
* Time period: 2010 - 2019
* "Once affordable": real estate tax in 2010 is below median value
* "Becoming unaffordable": change of real estate tax between 2010 and 2019 is greater than the median value

Then we can get these three categories: Gentrifying, Not-gentrifying, and Not-gentrifiable

|                     | **Significant RE Tax Increase (2010-2019)** | **No Significant RE Tax Change/Decrease (2010-2019)** |
|---------------------|:----------------------------------------:|:-----------------------------:|
| **Low RE Tax (2010)**   | Gentrifying                            | Not-gentrifying             |
| **High RE Tax (2010)**  | Gentrified                             | Gentrified                  |


## Data

We need: 

* Census ACS "median real estate tax" data for 2010 and 2019 in Fulton and Dekalb Counties
* Yoga studios within the two counties (will be provided)

### Let's first download the ACS 5-year estimate data for 2010 and 2019 for the two counties at the Tract level

```{r, results="hide"}
tract_2010 <- suppressMessages(
  get_acs(geography = ????,
          state = ????,
          county = c(????),
          variables = c(med_re_tax = "B25103_001"),
          year = 2010,
          survey = ????, 
          geometry = TRUE,
          output = ????)) %>% 
  select(GEOID, med_re_tax_2010 = med_re_taxE)

tract_2019 <- suppressMessages(
  get_acs(geography = ????,
          state = ????
          county = c(????),
          variables = c(med_re_tax = "B25103_001"),
          year = 2019,
          survey = ????, 
          geometry = FALSE, # why False?
          output = ????)) %>% 
  select(GEOID, med_re_tax_2019 = med_re_taxE)
```

<br>

### Join the two data and calculate the percentage change of median real estate tax between 2010 and 2019.
```{r}
tract <- left_join(tract_2010, 
                   tract_2019, 
                   by = "GEOID") %>% 
  mutate(med_re_tax_change = (med_re_tax_2019 - med_re_tax_2010)/med_re_tax_2010) %>% 
  drop_na(med_re_tax_change)

```

<br>

### Let's classify Census Tracts by the three categories we defined.
```{r}
print(paste0("Median value of median real estate tax in 2019: $", 
             median(tract$med_re_tax_2010) %>% round(2)))


print(paste0("Median value of the percentage change of median real estate tax between 2010 and 2019: ", 
             median(tract$med_re_tax_change) %>% round(2)))

tract <- tract %>% 
  mutate(gentri_category = case_when(
    med_re_tax_2010 >= median(med_re_tax_2010) ~ "Gentrified",
    med_re_tax_change >= median(med_re_tax_change) ~ "Gentrifying",
    TRUE ~ "Not-gentrifying"))

print(table(tract$gentri_category))
```

<br>

### Let's display what we have!

```{r, message=FALSE}
## tmap mode set to interactive viewing
tmap_mode("view")

tm_shape(tract) + tm_polygons(col = "med_re_tax_2019")
tm_shape(tract) + tm_polygons(col = "med_re_tax_change")
tm_shape(tract) + tm_polygons(col = "gentri_category")

```

<br>

### Now download the Yelp data. This has been done for you.

**To save time in class, we will use the Yelp data that is already downloaded for today's class.**

This data is for Fulton and DeKalb County, GA, and contains Yelp data with categories = "yoga". This data is already cleaned. To read the data into R, we will use st_read() function in sf package.

```{r}
# Reading the yelp data
yelp_yoga <- st_read("https://raw.githubusercontent.com/ujhwang/urban-analytics-2024/main/Lab/module_2/week1/yelp_yoga.geojson")
```

<br>

### Let's see if the results look fine

Get a summary using `skimr::skim()` and then display it (i.e., the point locations of yoga studios) on the Census Tract map.

```{r, warning=F}
skim(yelp_yoga)

tm_shape(yelp_yoga) + tm_dots(col="red") + tm_shape(tract) + tm_borders()
```

## Appending Census data

Joining census data in this case is somewhat different from what we've previously done. We will do the following:

1. make sure the CRS of the two sf objects are same
2. Calculate how many yoga studios are within each census tract (there will be many tracts with 0 studios) 
3. Join the two files with the count of Yoga studios

```{r, warning=F}
## Check to see that the CRS for both the sf files are the same
head(tract$geometry)
head(yelp_yoga$geometry)

## It turns out they are NOT! So we need to transform their CRS into the same
tract <- tract %>% st_transform(crs=4326) # you can skip 'crs='
yelp_yoga <- yelp_yoga %>% st_transform(crs=4326)

# Join tract geometry with yoga studios
tract_yoga <- st_join(tract, yelp_yoga %>% mutate(count = 1))

# Calculate the count of yoga studios at the tract level
tract_yoga_count <- tract_yoga %>%
  group_by(GEOID) %>%
  summarise(count = sum(count, na.rm = T))

## Now we are ready to join back the counts of yoga studios to the Tract data
tract <- tract %>%
  left_join(tract_yoga_count %>% st_drop_geometry(), 
            by = "GEOID")
```

### Make sure that the final dataset we will be using for our analysis is as expected

Let's skim and map the data

```{r}
????(tract)

# Let's check to see if the points of yoga studios and their counts in the polygon data match
tm_shape(????) + tm_polygons(col=????) +
  tm_shape(????) + tm_dots()
```

## Now we are ready to ask some probing questions about the data

### Are yoga studios more likely be in neighborhoods with ***high*** median real estate tax (2019)?

```{r}
ggplot(tract, aes(x=????, y=count)) +
  geom_point() +
  ylab("Number of Yoga Studios") +
  xlab("Median Real Estate Tax in Tract")
```

> Not much of an association...

<br>

### Are yoga studios more likely be in neighborhoods with ***increasing*** median real estate tax?

```{r}
ggplot(tract, aes(x=????, y=count)) +
  geom_point() +
  ylab("Number of Yoga Studios") +
  xlab("% Change of Real Estate Tax (2010-2019) in Tract")
```

> Hmm, it's showing some trend?

<br>

### Are yoga studios in **gentrifying neighborhoods**?

```{r}
tm_shape(tract) + tm_polygons(col=????) +
  tm_shape(yelp_yoga) +tm_dots()
```

> Most of yoga studios are in neighborhoods that are already gentrified.

<br>

A good way to quantify what we intuitively grasped from the previous map is to create a "boxplot"

```{r}
boxplot(count ~ gentri_category, data=tract, 
        main="Boxplot of Yoga Studios by Neighborhood Type", 
        xlab="Neighborhood types", 
        ylab="Yoga studio count")
```

> Gentrifying neighborhoods have more yoga studios than not-gentrifying neighborhoods, but less than already gentrified neighborhoods.

<br>

Let's convert it into a binary case: whether a Tract has a yoga studio or not.

```{r}
tract <- tract %>% 
  mutate(yoga_binary = count > 0)

yoga_binary_summary <- tract %>% 
  st_drop_geometry() %>% 
  group_by(yoga_binary, gentri_category) %>% 
  summarize(n = n()) %>% 
  ungroup() %>% 
  mutate(proportion = n / sum(n))

yoga_binary_summary %>% 
  ggplot(aes(x = gentri_category, y = proportion, fill = yoga_binary)) +
  geom_bar(stat = 'identity', position = 'fill')
```

> A similar story to the previous box plot.

## Finally, let's run a regression model.

* Dependent variable: **count of yoga studios**
* Independent variable: **type of neighborhood in terms of gentrification**
* Control variables:
  * **Educational attainment** (Rationale: higher education levels may correlate with interest in yoga)
  * **Median household income** (Rationale: income levels can influence the presence of amenities like yoga studios)
  * **Median age** (Rationale: Age demographics could influence the demand for yoga studios)
  * **Population density** (Rationale: denser areas might have more yoga studios due to higher foot traffic)
  * **Racial/ethnic composition** (Rationale: demographic makeup might correlate with yoga studio presence)
  * **% of population who commute by walking** (Rationale: it indicates pedestrian-friendly areas which may be more likely to have yoga studios)

Let's get these control variables.

```{r}
tract_control_vars <- suppressMessages(
  get_acs(geography = ????,
          state = ????,
          county = c(????),
          variables = c(edu_bachelor = "B06009_005",
                        edu_graduate = "B06009_006",
                        hhincome = "B19013_001",
                        median_age = "B01002_001",
                        pop = "B01003_001",
                        non_hispanic_white = "B03002_003",
                        commute_walk = "B08101_033",
                        commute_total = "B08101_001"),
          year = 2019,
          survey = ????, 
          geometry = TRUE,
          output = ????)) %>% 
  mutate(high_edu_pct = (edu_bachelorE + edu_graduateE)/popE,
         hhincome = hhincomeE,
         median_age = median_ageE,
         minority_pct = (popE - non_hispanic_whiteE)/popE,
         commute_walk_pct = commute_walkE/commute_totalE,
         pop = popE)

# Calculate land area of Census Tract and get population density using the area
tract_control_vars <- tract_control_vars %>% 
  mutate(area = st_area(.) %>% set_units(mi^2)) %>%
  mutate(pop_density = pop/area)

# Remove unnecessary columns including geometry
tract_control_vars <- tract_control_vars %>% 
  select(GEOID, high_edu_pct, hhincome, median_age, minority_pct, commute_walk_pct, pop_density) %>% 
  st_drop_geometry()

# Join the control variables to the original Tract data
tract <- tract %>% 
  left_join(tract_control_vars, by = ????)
```

<br>

Convert the `gentri_category` into a factor column where "Gentrifying" comes first in the category. (I will explain why)
```{r}
tract <- tract %>% 
  mutate(gentri_category = factor(gentri_category, 
                                  levels = c("Gentrifying",
                                             "Not-gentrifying",
                                             "Gentrified")))
```

<br>

Run multiple linear regression.
```{r}
m1 <- lm(count ~ gentri_category + high_edu_pct + hhincome + 
            median_age + minority_pct + commute_walk_pct + pop_density, 
          data=tract)
summary(m1)
```

> How would you interpret this result?

### So what do you think about our initial question? ***Are yoga studios associated with gentrification?***

### Would the result be different if we use ***median rent*** instead of *median real estate tax*?
