---
title: "Major_assignment_1_template"
author: "Originally written by Bon Woo Koo & Subhro Guhathakurta; modified by Uijeong Hwang"
date: '2024-10-16'
output: html_document
---



# How to use this template

You will see # TASK ///// through out this template. This indicates the beginning of a task.  Right below it will be instructions for the task.
Each # TASK ///// will be paired with # //TASK ///// to indicate where that specific task ends.

For example, if you see something like below...
```{r example}
# TASK ////////////////////////////////////////////////////////////////////////
# create a vector with element 1,2,3 and assign it into `my_vec` object
# **YOUR CODE HERE..**
# //TASK //////////////////////////////////////////////////////////////////////
```

What I expect you to do is to replace where it says `# **YOUR CODE HERE..**` with your answer, like below.
```{r}
# TASK ////////////////////////////////////////////////////////////////////////
# create a vector with element 1,2,3 and assign it into `my_vec` object
my_vec <- c(1,2,3)
# //TASK //////////////////////////////////////////////////////////////////////
```


There can be multi-step instructions, like shown below. You may use pipe (%>%) to link multiple functions to perform the task in the instruction. Make sure that **you assign the output of your task into an object with the specified name**. This is to make sure that your code will run smoothly - if you change the name of the object (i.e., subset_car in the example below), all the subsequent code will NOT run properly.
```{r}
# TASK ////////////////////////////////////////////////////////////////////////
# 1. Using mtcars object, extract rows where cyl equals 4
# 2. Select mpg and disp columns
# 3. Create a new column 'summation' by adding mpg and disp
# 4. assign it into `subset_car` object
subset_car <- # **YOUR CODE HERE..**
# //TASK //////////////////////////////////////////////////////////////////////
```

I expect you to replace where it says `# **YOUR CODE HERE..**` with your answer, like below.
```{r}
# TASK ////////////////////////////////////////////////////////////////////////
# 1. Using mtcars object, extract rows where cyl equals 4
# 2. Select mpg and disp columns
# 3. Create a new column 'summation' by adding mpg and disp
# 4. assign it into `subset_car` object
subset_car <- mtcars %>% 
  filter(cyl == 4) %>% 
  select(mpg, disp) %>% 
  mutate(summation = mpg + disp)
# //TASK //////////////////////////////////////////////////////////////////////
```

> You will need to knit it, publish it on Rpubs, and submit the link.

# Task description

There are a few main components in this assignment - home location, road networks, transit network, and destination. We will simulate a journey that starts from the starting point (e.g., home), drives to the nearest MARTA rail station, transfers to MARTA rail transit, and finally arrives at Midtown station. The following is a list of tasks and data we need for this analysis.

Step 1. Download Required data from GTFS. Convert it to sf format, extract MARTA rail stations, and clean the stop names to delete duplicate names. Also extract the destination station.

Step 2. Download Required data from Census. Convert Census polygons into centroids and create a subset.

Step 3. Download Required data from OSM. Convert it into an sfnetwork object and clean the network.

Step 4. Simulate a park-and-ride trip (home -> closest station -> Midtown station).

Step 5. Convert what we did in Step 4 into a function so that we can use it to repeat it in a loop.

Step 6. Run a loop to repeat the function from Step 5 to all other home location. Once finished, merge the simulation output back to Census data.

Step 7. Finally, examine whether there is any disparity in using transit to commute to midtown.

Before we start, libraries first..

```{r}
library(tidyverse)
library(tmap)
library(units)
library(sf)
library(leaflet)
library(dbscan)
library(sfnetworks)
library(tigris)
library(tidygraph)
library(plotly)
library(osmdata)
library(here)
library(tidytransit)
library(tidycensus)
library(leafsync)

epsg <- 4326
```

# Step 1. Download Required data from GTFS.

```{r}
# TASK ////////////////////////////////////////////////////////////////////////
# Download GTFS data using `read_gtfs()` function and assign it in `gtfs` object
gtfs <- # **YOUR CODE HERE..**
# //TASK //////////////////////////////////////////////////////////////////////



# =========== NO MODIFICATION ZONE STARTS HERE ===============================
# Edit stop_name to append serial numbers (1, 2, etc.) to remove duplicate names
stop_dist <- stop_group_distances(gtfs$stops, by='stop_name') %>%
  filter(dist_max > 200)

gtfs$stops <- gtfs$stops %>% 
  group_by(stop_name) %>% 
  mutate(stop_name = case_when(stop_name %in% stop_dist$stop_name ~ paste0(stop_name, " (", seq(1,n()), ")"),
                               TRUE ~ stop_name))

# Create a transfer table
gtfs <- gtfsrouter::gtfs_transfer_table(gtfs, 
                                        d_limit = 200, 
                                        min_transfer_time = 120)

# NOTE: Converting to sf format uses stop_lat and stop_lon columns contained in gtfs$stops.
#       In the conversion process, stop_lat and stop_lon are converted into a geometry column, and
#       the output sf object do not have the lat lon column anymore.
#       But many other functions in tidytransit look for stop_lat and stop_lon.
#       So I re-create them using mutate().
gtfs <- gtfs %>% gtfs_as_sf(crs = epsg)

gtfs$stops <- gtfs$stops %>% 
  ungroup() %>% 
  mutate(stop_lat = st_coordinates(.)[,2],
         stop_lon = st_coordinates(.)[,1]) 

# Get stop_id for rails and buses
rail_stops <- gtfs$routes %>% 
  filter(route_type %in% c(1)) %>% 
  inner_join(gtfs$trips, by = "route_id") %>% 
  inner_join(gtfs$stop_times, by = "trip_id") %>% 
  inner_join(gtfs$stops, by = "stop_id") %>% 
  group_by(stop_id) %>% 
  slice(1) %>% 
  pull(stop_id)

# Extract MARTA rail stations
station <- gtfs$stops %>% filter(stop_id %in% rail_stops)

# Extract Midtown Station
midtown <- gtfs$stops %>% filter(stop_id == "134")

# Create a bounding box to which we limit our analysis
bbox <- st_bbox(c(xmin = -84.45241, ymin = 33.72109, xmax = -84.35009, ymax = 33.80101), 
                 crs = st_crs(4326)) %>% 
  st_as_sfc()

# =========== NO MODIFY ZONE ENDS HERE ========================================

```

# Step 2. Download Required data from Census

```{r}
# TASK ////////////////////////////////////////////////////////////////////////
# Specify Census API key whichever you prefer using census_api_key() function
census_api_key(" **YOUR CODE HERE..** ")
# //TASK //////////////////////////////////////////////////////////////////////



# TASK ////////////////////////////////////////////////////////////////////////
# Using get_acs() function, download Census Tract level data for 2022 for Fulton, DeKalb, and Clayton in GA.
# and assign it to `census` object.
# Make sure you set geometry = TRUE.

# Required data from the Census ACS:
#  1) Median Household Income (name the column `hhinc`)
#  2) Minority Population (%) (name the column `pct_minority`)
# Note: You may need to download two or more Census ACS variables to calculate minority population (%). "Minority" here can refer to either racial minorities or racial+ethnic minorities -- it's your choice.

census <- # **YOUR CODE HERE..**

# //TASK //////////////////////////////////////////////////////////////////////



# =========== NO MODIFICATION ZONE STARTS HERE ===============================
census <- census %>% 
  st_transform(crs = 4326) %>% 
  separate(col = NAME, into = c("tract", "county", "state"), sep = ", ")

# Convert it to POINT at polygon centroids and extract those that fall into bbox
# and assign it into `home` object
home <- census %>% st_centroid() %>% .[bbox,]
# =========== NO MODIFY ZONE ENDS HERE ========================================

```



# Step 3. Download Required data from OSM.

```{r}
# TASK ////////////////////////////////////////////////////////////////////////
# 1. Get OSM data using opq() function and bbox object defined in the previous code chunk.
# 2. Specify arguments for add_osm_feature() function using 
#    key = 'highway' and 
#    value = c("motorway", "trunk", "primary", "secondary", "tertiary", "residential", 
#              "motorway_link", "trunk_link", "primary_link", "secondary_link", 
#              "tertiary_link", "residential_link", "unclassified")
# 3. Convert the OSM data into an sf object using osmdata_sf() function
# 4. Convert osmdata polygons into lines using osm_poly2line() function

osm_road <- # **YOUR CODE HERE..**
# //TASK //////////////////////////////////////////////////////////////////////


# TASK ////////////////////////////////////////////////////////////////////////
# 1. Convert osm_road$osm_lines into sfnetworks using as_sfnetwork() function
# 2. Activate edges
# 3. Clean the network using edge_is_multiple(), edge_is_loop(), to_spatial_subdivision(), to_spatial_smooth()
# 4. Assign the cleaned network to an object named 'osm'

osm <- osm_road$osm_line %>% 
  # **YOUR CODE HERE..**
  # ...
# //TASK //////////////////////////////////////////////////////////////////////



# TASK ////////////////////////////////////////////////////////////////////////
# Add a new column named 'length' to the edges part of the object `osm`.
osm <- osm %>% 
  # **YOUR CODE HERE..**
  # ...
# //TASK //////////////////////////////////////////////////////////////////////
```

# Step 4. Simulate a park-and-ride trip (home -> closest station -> Midtown station).

```{r}

# =========== NO MODIFICATION ZONE STARTS HERE ===============================
# Extract the first row from `home` object and store it `home_1`
home_1 <- home[1,]
# =========== NO MODIFY ZONE ENDS HERE ========================================


# TASK ////////////////////////////////////////////////////////////////////////
# Find the shortest path from `home_1` to all other stations
# using st_network_paths() function.
paths <- 
# //TASK //////////////////////////////////////////////////////////////////////

  
# =========== NO MODIFICATION ZONE STARTS HERE ===============================
# Using the `paths` object, get network distances from `home_1` to all other stations.
dist_all <- map_dbl(1:nrow(paths), function(x){
  atl_net %>% 
    activate("nodes") %>% 
    slice(paths$node_paths[[x]]) %>% 
    st_as_sf("edges") %>% 
    pull(length) %>% 
    sum()
}) %>% unlist() 

# Replace zeros with a large value.
if (any(dist_all == 0)){
  dist_all[dist_all == 0] <- max(dist_all)
}

# Find the index of the closest station.
closest_index <- which.min(dist_all)

# Find the distance to the closest station.
closest_dist <- min(dist_all)

# Calculate how long it takes to traverse `closest_dist` 
# assuming we drive at 30 miles/hour speed.
# Store the output in trvt_osm_m.
car_speed <- set_units(30, mile/h)
trvt_osm_m <- closest_dist/set_units(car_speed, m/min) %>%  # Distance divided by 30 mile/h
  as.vector(.)
# =========== NO MODIFY ZONE ENDS HERE ========================================


# TASK ////////////////////////////////////////////////////////////////////////
# 1. From `osm` object, activate nodes part and
# 2. use `closest_index` to extract the selected path
paths_closest <- # **YOUR CODE HERE..**
# //TASK //////////////////////////////////////////////////////////////////////


# TASK ////////////////////////////////////////////////////////////////////////
# Use filter_stop_times() function to create a subset of stop_times data table
# for date = 2021-08-14, minimum departure time of 7AM, maximum departure time of 10AM.
# Assign the output to `am_stop_time` object
am_stop_time <- # **YOUR CODE HERE..**
# //TASK //////////////////////////////////////////////////////////////////////



# TASK ////////////////////////////////////////////////////////////////////////
# 1. Use travel_times() function to calculate travel times from the `closest_station` 
#    to all other stations during time specified in am_stop_time. 
# 2. Filter the row for which the value of 'to_stop_name' column 
#    equals midtown$stop_name. Assign it into `trvt` object.
trvt <-  # **YOUR CODE HERE..**
# //TASK //////////////////////////////////////////////////////////////////////



# =========== NO MODIFICATION ZONE STARTS HERE ===============================
# Divide the calculated travel time by 60 to convert the unit from seconds to minutes.
trvt_gtfs_m <- trvt$travel_time/60

# Add the travel time from home to the nearest station and
# the travel time from the nearest station to Midtown station
total_trvt <- drop_units(trvt_osm_m) + trvt_gtfs_m
# =========== NO MODIFY ZONE ENDS HERE ========================================
```


# Step 5. Convert Step 4 into a function

```{r}
# Function definition (do not modify other parts of the code in this code chunk except for those inside the TASK section)

get_trvt <- function(home, osm, station, midtown){
  
  # TASK ////////////////////////////////////////
  # If the code in Step 4 runs fine,
  # Replace where it says **YOUR CODE HERE..** below with 
  # the ENTIRETY of the code in the previous code chunk (i.e., Step 4)
  
  # **YOUR CODE HERE..**
  
  # //TASK //////////////////////////////////////

  # =========== NO MODIFICATION ZONE STARTS HERE ===============================
  if (length(total_trvt) == 0) {total_trvt = 0}

  return(total_trvt)
  # =========== NO MODIFY ZONE ENDS HERE ========================================
}
```


# Step 6. Apply the function for the whole study area

```{r}
# Prepare an empty vector
total_trvt <- vector("numeric", nrow(home))

# Apply the function for all Census Tracts
# Fill `total_trvt` object with the calculated time
for (i in 1:nrow(home)){
  total_trvt[i] <- get_trvt(home[i,], osm, station, midtown)
}

# Cbind the calculated travel time back to `home`
home_done <- home %>% 
  cbind(trvt = total_trvt)
```


# Step 7. Create maps and plots

**Run the code below to generate thematic maps and plots**

**Write a short description of what you observe from the maps and plots**

```{r}
# Map
tmap_mode('view')

tm_shape(census[census$GEOID %in% home$GEOID,]) + 
  tm_polygons(col = "hhinc", palette = 'GnBu') + 
  tm_shape(home_done) + 
  tm_dots(col = "trvt", palette = 'Reds', size = 0.1)

tm_shape(census[census$GEOID %in% home$GEOID,]) + 
  tm_polygons(col = "pct_minority", palette = 'GnBu') + 
  tm_shape(home_done) + 
  tm_dots(col = "trvt", palette = 'Reds', size = 0.1)

# ggplot
inc <- ggplot(data = home_done,
              aes(x = hhinc, y = trvt)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Median Annual Household Income",
       y = "Park-and-ride Travel Time from Home to Midtown Station") +
  theme_bw()

minority <- ggplot(data = home_done,
                   aes(x = pct_minority, y = trvt)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Minority Population (%)",
       y = "Park-and-ride Travel Time from Home to Midtown Station") +
  theme_bw()


ggpubr::ggarrange(inc, wh)
```

