---
title: "Visualization for data exploration & communication"
author: "Originally written by Bon Woo Koo & Subhro Guhathakurta; modified by Uijeong Hwang"
date: '2024-09-30'
output:
  rmdformats::downcute:
    downcute_theme: "chaos"
---

```{=html}
<style type="text/css">
  body{
  font-family: Arial;
  }
</style>
```
# Why visualization?

Data visualization is an essential tool for urban analytics. It offers an intuitive and quick way to understand the characteristics and relationships embedded in your data. Sometimes a good data visualization can deliver a message more effectively than any other medium and exercise transformative power that shapes people's perception [see this example](https://www.cnn.com/interactive/2021/us/census-race-ethnicity-map/). And not to mention that cool visualizations are cool.

The first thing we analysts do when we get our hands on a dataset is to understand it. The best way to understand any given data is to visualize it. We have a wide variety of maps and charts to choose from, including scatter plot, histogram, box plot, violin plot, and mapping. We have been doing mapping a lot, so this document will focus more on plots.

```{r, message=FALSE, results='hide', warning=FALSE}
library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
library(tidycensus)
```

```{r, results='hide', message=FALSE, warning=FALSE}
# Let's prepare data
yelp <- read_rds("https://raw.githubusercontent.com/ujhwang/urban-analytics-2024/main/Lab/module_2/week2/yelp_in.rds")

# Census data
census_var <- c(hhincome = 'B19019_001',
                pop = "B02001_001",
                race.white = "B02001_002", 
                race.black = 'B02001_003'
                )

census <- get_acs(geography = "tract", state = "GA", county = c("Fulton", "DeKalb"),
                 output = "wide", geometry = TRUE, year = 2020,
                 variables = census_var)
  
summarise_mean <- c(str_c(names(census_var), "E"), 
                    "rating", "review_count")

census_yelp <- census %>% 
  separate(col = NAME, into=c("tract","county","state"), sep=', ') %>% 
  # Spatial join
  st_join(yelp %>% 
            mutate(n = 1,
                   price = nchar(price)) %>% 
            st_transform(crs = st_crs(census))) %>% 
  # Group_by
  group_by(GEOID, county) %>% 
  # Mean for all census variables, sum for n
  summarise(across(
    all_of(summarise_mean), mean), 
    n = sum(n),
    price = median(price)) %>% 
  # Release grouping
  ungroup() %>% 
  # Drop 'E' from column names
  rename_with(function(x) str_sub(x,1,nchar(x)-1), str_c(names(census_var), "E")) %>% # rename_with() renames with a function
  # Replace NA in column n&review_count with 0
  mutate(across(c(n, review_count), function(x) case_when(is.na(x) ~ 0, TRUE ~ x)))
```

As usual, using tmap to visualize the data.

```{r warning=F}
tmap_mode("view");
a <- tm_shape(census_yelp) + 
  tm_????(col = "review_count", style = "quantile")

b <- tm_shape(yelp) +
  tm_????(col = "review_count", style="quantile")

tmap_arrange(a,b, sync = TRUE)
```

Or we can use leaflet() package for mapping.

```{r warning=F}
 library(leaflet)
 library(htmlwidgets)
 library(htmltools)

# CSS for title
tag.map.title <- tags$style(HTML("
  .leaflet-control.map_title { 
    position: absolute;
    left: 50px;
    width: 320px;
    text-align: left;
    color: white;
    padding-left: 10px; 
    background: rgba(200,200,200,0.2);
    font-weight: bold;
    font-size: 20px;
    font-family: Helvetica;
    border-color: white;
    border-radius: 10px;
  }"))

# Format title
title <- tags$div(
  class="map_title", tag.map.title, HTML("<p>Restaurants in Fulton and DeKalb Counties from Yelp</p>")
)  

# Color palette
fill_pal <- colorQuantile(palette = "YlOrRd", domain=yelp$review_count)

# Label for mouseover & popup
yelp_labels <- paste(
  "<a href=",yelp$url,">",yelp$name, "</a><br>",
  "<strong>Review Count: </strong>", yelp$review_count,"<br>",
  "<strong>Rating: </strong>", yelp$rating) %>%
  lapply(htmltools::HTML)

# Creating a Leaflet widget
leaflet() %>% 
  # Setting the view on load
  setView(lng = -84.3903996350635, lat = 33.77074368998939, zoom = 11) %>% 
  
  # Dark base map
  addProviderTiles(providers$CartoDB.DarkMatterNoLabels) %>% 
  
  # Polygon boundary
  addPolygons(data = census %>% st_union(), 
              opacity=0.2,
              fillOpacity=0,
              weight=1,
              color="white") %>% 
  
  # Yelp point
  addCircleMarkers(data = yelp, 
             radius = yelp$rating*1.5, 
             opacity=0.2,
             fillColor=~fill_pal(review_count),
             weight=1,
             color= ~fill_pal(review_count),
             popup= ~yelp_labels,
             label= ~yelp_labels) %>% 
  
  # Legend
  addLegend("bottomright", pal = fill_pal, values = yelp$review_count,title = "Review Count",opacity = 1) %>% 
  
  # Title
  addControl(title, position="topleft",className="map_title")
```

<br><br> For static maps with a lot of customizeability - ggplot2 can also create maps. See [this example](https://github.com/grssnbchr/bivariate-maps-ggplot2-sf) for an inspiration.

```{r}
library(viridis)

ggplot(data = yelp) +
  geom_point(mapping = aes(color = log(review_count+1), 
                           x = st_coordinates(yelp)[,1], 
                           y = st_coordinates(yelp)[,2]), 
             size=2, alpha=0.9) +
  theme_void() +
  scale_color_viridis(trans = "log",  
                      name="Review Count (logged)", 
                      guide = guide_legend(keyheight = unit(3, units = "mm"),
                                           keywidth=unit(12, units = "mm"), 
                                           label.position = "bottom", 
                                           title.position = 'top', nrow=1)) +
  labs(
    title = "Restaurant Distribution in Fulton & DeKalb Counties, GA",
    subtitle = "Review Count from Yelp API"
  ) +
  theme(
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#f5f5f2", color = NA),
    panel.background = element_rect(fill = "#f5f5f2", color = NA),
    legend.background = element_rect(fill = "#f5f5f2", color = NA),

    plot.title = element_text(size= 12, color = "black", 
                              margin = margin(b = -0.1, t = 0.4, l = 2, unit = "cm")),
    plot.subtitle = element_text(size= 10, color = "black", 
                                 margin = margin(b = -0.1, t = 0.43, l = 2, unit = "cm")),
  plot.caption = element_text( size=7, hjust=0, color = "black", 
                                 margin = margin(b = 0.3, r=-99, unit = "cm") )
  ) +
  coord_map()

```

# Scatter plot

In tmap package, two functions always go hand-in-hand, namely tm_shape() and tm_polygons() (or tm_lines, tm_dots, etc.). The tm_shape() function declares the data object to be displayed. Then, the tm_polygons() function defines the geometry shape and other associated characteristics.

A similar structure is used in ggplot2 package. Creating a plot needs at least two functions that are connected by +: ggplot() function and geom_point() (or other geometry types, such as geom_line, geom_box plot, etc.). In the example below, `ggplot(data = census_yelp)` indicates that we are drawing ggplot using yelp data. Then, `geom_point(aes(x = review_count, y = rating))` shows that we are going to draw a scatter plot using `review_count` and `rating` columns.

```{r warning=F}
ggplot(data = ????) +
  geom_point(mapping = aes(x = ????, y = ????))
```

## Aesthetic mappings

We can add additional information to this plot using a few different strategies, including colors, sizes, and shapes inside the `aes()` -- 'Aesthetic mappings' which describe how variables in the data are mapped to visual properties (aesthetics) of geoms. For example, we can add `price` information using color.

```{r warning=F}
ggplot(data = ????) +
  geom_point(mapping = aes(x=????, y=????, 
                           color=????)) #<<
```

Or size.

```{r warning=F}
ggplot(data = ????) +
  geom_point(mapping = aes(x=????, y=????, 
                          size=????)) #<<
```

Or transparency.

```{r warning=F}
fig1 <- ggplot(data = ????) +
  geom_point(mapping = aes(x=????, y=????, 
                           alpha=price)) #<<

fig2 <- ggplot(data = ????) +
  geom_point(mapping = aes(x=????, y=????, 
                           alpha=price, #<<
                           size=price)) #<<

gridExtra::grid.arrange(????, ????, ncol= 1)
```

If you put, for example, color and size arguments outside of `aes()`, the visual property from those arguments are not based on your data.

```{r warning=F}
fig2 <- ggplot(data = census_yelp) +
  geom_point(mapping = aes(x=review_count, y=rating), 
             color = "orange") #<<

fig3 <- ggplot(data = census_yelp) +
  geom_point(mapping = aes(x=review_count, y=rating), 
             size = 5) #<<

gridExtra::grid.arrange(fig2, fig3, ncol = 1)
```

## Separating plots by categorical values

Use `facet_wrap` with a categorical variable. Remember that if you use `facet_wrap` with a continuous data, it will generate as many plots as the unique values in the continuous data. Avoid writing such codes!

```{r warning=F}
ggplot(data = census_yelp) +
  geom_point(mapping = aes(x=review_count, y=rating)) +
  facet_wrap(~county) #<<
```

```{r warning=F}
ggplot(data = census_yelp) +
  geom_point(mapping = aes(x=review_count, y=rating)) +
  facet_grid(???? ~ ????) #<<
# One will be rows; the other will be columns.
# try swapping the two
```

## Regression line

You can draw a regression line by using `geom_smooth` with the "lm" method.

```{r warning=F}
ggplot(data = census_yelp) +
  geom_????(mapping = aes(x=review_count, y=rating), method = ????)
```

More informative if you show both data points and the regression line.

```{r warning=F}
# More than one layers
ggplot(data = census_yelp) +
  geom_point(mapping = aes(x=review_count, y=rating)) + 
  geom_????(mapping = aes(x=review_count, y=rating), method = ????)
```

In the code above, we are repeating `aes(x=review_count, y=rating)` twice. If we know that mapping will be the same in multiple layers, we can define it in `ggplot()`.

```{r warning=F}
# You don't need to repeat `mapping = aes(x=review_count, y=rating)` twice
ggplot(data = census_yelp, mapping = aes(x=review_count, y=rating)) + #<<
  geom_point() + 
  geom_????(method = ????)
```

If you want to add a specific mapping to a layer, you provide additional mapping to individual layers.

```{r warning=F}
ggplot(data = census_yelp, mapping = aes(x=review_count, y=rating)) +
  geom_point() + 
  geom_????(mapping = aes(color = ????), #<<
            method = ????)
```


You can append `labs()` to specify labels.

```{r warning=F}
ggplot(data = census_yelp, 
       ???? = aes(x=review_count, y=rating)) +
  geom_point() + 
  geom_????(???? = aes(color = ????), method = ????) +
  labs(x = "Review Count in Yelp", #<<
       y = "Rating in Yelp",
       color = "County in Census",
       title = "Do better rated restaurants have more reviews?")
```

# Aesthetic options

## Theme

We can change the overall theme of the plot using `theme_<...>`.

```{r warning=F}
ggplot(data = census_yelp) +
  geom_point(mapping = aes(x=review_count, y=rating, color = county)) + 
  labs(x = "Review Count in Yelp",
       y = "Rating in Yelp",
       color = "County in Census",
       title = "Do better rated restaurants have more reviews?") +
  theme_bw()
```

Of course, dark is always cooler.

```{r warning=F}
ggplot(data = census_yelp) +
  geom_point(mapping = aes(x=review_count, y=rating, color = county)) + 
  labs(x = "Review Count in Yelp",
       y = "Rating in Yelp",
       color = "County in Census",
       title = "Do better rated restaurants have more reviews?") +
  ggdark::dark_theme_gray()
  
```

## Custom color scheme

If you want to use your custom color choices - for a discrete variable.

```{r warning=F}
ggplot(data = census_yelp) +
  geom_point(mapping = aes(x=review_count, y=rating, color = county)) + 
  labs(x = "Review Count in Yelp",
       y = "Rating in Yelp",
       color = "County in Census",
       title = "Do better rated restaurants have more reviews?") +
  scale_color_manual(values = c("green", "darkblue")) + #<<
  theme_bw()
```

If you want to use your custom color choices - for a continuous variable.

```{r warning=F}
ggplot(data = census_yelp) +
  geom_point(mapping = aes(x=review_count, y=rating, color = hhincome)) + 
  labs(x = "Review Count in Yelp",
       y = "Rating in Yelp",
       color = "Annual Household Income",
       title = "Do better rated restaurants have more reviews?") +
  scale_color_gradient(low="darkblue", high="red") + #<<
  theme_bw()
```

## Data labeling

Let's label some of the outliers using `ggrepel`.

```{r warning=F}
outliers <- census_yelp %>% 
  arrange(desc(review_count)) %>% 
  slice(1:4)

ggplot(data = census_yelp,
       aes(x=review_count, y=rating)) + # moved aes() to here
  geom_point(mapping = aes(color = ????)) + # Colored ones
  geom_point(data = outliers, size = 3, shape = 1, color = "black") + # Black circles
  ggrepel::geom_label_repel(data = census_yelp, mapping = aes(label = ????)) + 
  labs(x = "Review Count in Yelp",
       y = "Rating in Yelp",
       color = "Annual Household Income",
       title = "Do better rated restaurants have more reviews?") +
  scale_color_gradient(low="darkblue", high="red") + #<<
  theme_light()
```

## Interactive visualization

How can we convert these plots into interactive ones? The simplest way is to use `ggplotly()` in `plotly` package.

```{r warning=F}
p <- ggplot(data = census_yelp) +
  geom_point(mapping = aes(x=review_count, y=rating, color = hhincome)) + 
  labs(x = "Review Count in Yelp",
       y = "Rating in Yelp",
       color = "Annual Household Income",
       title = "Do better rated restaurants have more reviews?") +
  scale_color_gradient(low="darkblue", high="red") + #<<
  theme_bw()

plotly::????(p)
```

## 3D scatter plot

Using `plot_ly` in `plotly` package, we can also create a 3-dimensional scatter plot. It is particularly useful when you want to examine associations between four variables.

```{r}
plotly::????(census_yelp %>% st_drop_geometry(),
                x = ~hhincome,
                y = ~n,
                z = ~rating,
                color = ~price,
                type = "scatter3d",
                mode = "markers",
                text = ~paste0("Avg. price: ", round(price, 2)),
                hoverinfo = "text")
```

# Bar plot

Bar plot is also very frequently used. Note that ggplot creates y-axis automatically by examining how many rows there are for each category of x. You can try `yelp %>% group_by(price) %>% tally()` to check the exact Y value for this plot.

```{r}
ggplot(data = yelp) +
  geom_bar(mapping = aes(x=price)) 

yelp %>% 
  st_set_geometry(NULL) %>% 
  group_by(price) %>% 
  tally()
```

We can also further break each price level by another categorical variable. We use rating to see the relative frequency of each rating for each price level. This is done by adding `fill=rating` in the mapping.

```{r}
ggplot(data = yelp %>% 
         st_set_geometry(NULL) %>% 
         mutate(rating = round(rating,0) %>% #<<
                  factor(ordered = TRUE))) + #<<
         # delete %>% factor(ordered = T) and see what happens
  geom_bar(mapping = aes(x=????, fill=rating), position = "stack")
```

By changing `position="stack"` to `position="fill"`, we convert the Y-axis to the proportion within each level of price and fill it up to the top. This shows more clearly how different rating levels are distributed within each price level.

```{r}
ggplot(data = yelp %>% 
         st_set_geometry(NULL) %>% 
         mutate(rating = round(rating,0) %>% 
                  factor(ordered = TRUE))) + #<<
  geom_bar(mapping = ????(x=????, fill=rating), position = ????) #<<
```

We want rating=5 to be on the top because it is more intuitive to see higher value on top. We can flip the bar plot vertically by adjusting the `levels` when we declare `rating` variable into a factor.

```{r}
ggplot(data = yelp %>% 
         st_set_geometry(NULL) %>% 
         mutate(rating = round(rating,0) %>% 
                  factor(levels = c(5,4,3,2,1), #<<
                         ordered = TRUE))) +
  geom_bar(mapping = ????(x=????, fill=rating), position = ????)
```

Sometimes we want to see the exact figures on top of the bar plot. So, let's add the percentage within each level of price as labels.

```{r}
ggplot(data = yelp %>% 
         st_set_geometry(NULL) %>% 
         mutate(rating = round(rating,0) %>% factor(levels = c(5,4,3,2,1), ordered = TRUE))) +
  geom_bar(???? = aes(x=????, fill=rating), position = ????) +
  geom_text(data = . %>% 
              # Grouping to calculate % by price and by rating 
              group_by(price, rating) %>% #<<
              # Count rows
              tally() %>%                 #<< 
              # Convert to p
              mutate(p = n / sum(n)) %>%  #<<
              # Re-order to match the order in bar plot
              arrange(desc(rating)),      #<<
            aes(x = price, y = p, label = str_c(round(p,3)*100,"%")), color = "white",
            position = position_stack(vjust=0.5)) +
  ggdark::dark_theme_gray() # Dark theme because texts are not visible against white bg
```

You can flip it 90-degrees.

```{r}
ggplot(data = yelp %>% 
         st_set_geometry(NULL) %>% 
         mutate(rating = round(rating,0) %>% factor(levels = c(5,4,3,2,1), ordered = TRUE))) +
  geom_bar(???? = aes(x=????, fill=rating), position = ????) +
  geom_text(data = . %>% 
              # Grouping to calculate % by price and by rating 
              group_by(price, rating) %>%
              # Count rows
              tally() %>%
              # Convert to p
              mutate(p = n / sum(n)) %>%
              # Re-order to match the order in bar plot
              arrange(desc(rating)),
            aes(x = price, y = p, label = str_c(round(p,3)*100,"%")), color = "white",
            position = position_stack(vjust=0.5)) +
  coord_flip() + #<<
  ggdark::dark_theme_gray() # Dark theme because texts are not visible against white bg
```

## Customization example (Optional)

I saw this beautiful example by CÉDRIC SCHERER and wanted to show you a Yelp version of the code.

```{r}
# Code & ideas borrowed heavily from CÉDRIC SCHERER's personal website: 
# https://www.cedricscherer.com/2021/07/05/a-quick-how-to-on-labelling-bar-graphs-in-ggplot2/

max_city_n <- 10

rest_by_city <- yelp %>% 
  st_set_geometry(NULL) %>% 
  group_by(location.city) %>% 
  tally() %>% 
  arrange(desc(n)) %>% 
  slice(1:max_city_n) %>% 
  mutate(location.city = factor(location.city, levels = .$location.city[seq(max_city_n,1)])) %>% 
  # Format text label
  mutate(pct = scales::percent(n / sum(n),accuracy = 0.1),
         pct = case_when(row_number() == 1 ~ str_c(pct, " of all businesses"), TRUE ~ pct)) %>% 
  # Define aesthetic properties - label location
  mutate(nudge = case_when(row_number()==1 ~ 1.05, TRUE ~ -0.2)) %>% 
  # Define aesthetic properties - color
  mutate(color = case_when(row_number()==1 ~ "gray30", TRUE ~ "gray70")) %>% 
  # Color palette
  mutate(pal = c(rep('gray70', max_city_n-4), "coral2", "mediumpurple1", "mediumpurple1", "goldenrod1")) %>% 
  # with() is required to be able to call variables with referencing to data frame
  with(
    # ggplot
    ggplot(data = .) +
      # Bars
      geom_col(mapping = aes(y = location.city, x = n, fill = location.city)) +
      # Text
      geom_text(mapping = aes(y = location.city, x = n, label = pct), 
                # Calling aesthetic properties defined above
                hjust=nudge, color=color, 
                # Font styling
                fontface="bold.italic") + 
      # Stretch x axis
      scale_x_continuous(limits = c(NA, 2200)) +
      # Custom palette
      scale_fill_manual(values = pal, guide="none") +
      # Labels
      labs(x = "Count", y = "Cities", title = "Top 10 cities with most restaurants in Fulton & DeKalb Counties\n") +
      # Dark theme
      ggdark::dark_theme_classic()
  )

rest_by_city
```

# Histogram, box plot, & violin plot

Histogram displays the distribution of a variable. It first assigns observations (i.e., rows) of the given variable (i.e., a column) into bins (0\~99, 100\~199, 200\~299, etc.) and count the number of observations that fall into each bin. So, the taller a bar is, more observations fell into that bin. For example, in the histogram below, the first bar that touches 0 is much taller than other bars because most of the observations had zero or near-zero reviews.

Let's examine the distribution of review counts using histogram.
```{r}
ggplot(census_yelp) +
  geom_histogram(mapping = ????(x = ????))
```

```{r}
ggplot(census_yelp) +
  geom_histogram(mapping = ????(x = ????), 
                 bins = 60) #<< increase the number of bins from 30 to 50. You get more bars.
```

```{r}
ggplot(census_yelp) +
  geom_histogram(mapping = ????(x = ????), 
                 bins = 60,
                 color="black")  #<< color of the outline
```

```{r}
ggplot(data = census_yelp) +
  geom_histogram(mapping = ????(x = ????, fill=county), #<<
                 bins = 60,
                 color="black") +
  scale_x_continuous(breaks=seq(0,1300, by=100))
```

```{r}
ggplot(data = census_yelp) +
  geom_histogram(mapping = ????(x = ????, fill=county), 
                 bins = 60,
                 color="black",
                 position = "identity", #<<
                 alpha = 0.2) #<<
  scale_x_continuous(breaks=seq(0,1300, by=100))
```

```{r}
ggplot(data = census_yelp) +
  geom_histogram(mapping = ????(x = ????, fill=county), 
                 bins = 60,
                 color="black",
                 position = "dodge") + #<< #<<
  scale_x_continuous(breaks=seq(0,1300, by=100))
```

```{r}
ggplot(data = census_yelp) +
  geom_histogram(mapping = ????(x = ????, fill=county), 
                 bins = 60,
                 color="black",
                 position = "dodge") + #<< #<<
  scale_x_continuous(breaks=seq(0,1300, by=100)) + 
  scale_fill_manual(values=c("#999999", "#E69F00"))
```

Although the histogram above allows us to effectively see the distribution, comparing **multiple distributions** are often easier with a box plot. You need to understand what each component of a box plot means to read it properly. `plotly::ggplotly()` provides a good interactive visualization about how to read the plot. Note that upper fence = Q3 + (1.5\*IQR), where IQR is interquartile range (Q3-Q1). The lower fence is Q1 - (1.5\*IQR).

```{r}
bxplot <- ggplot(data = yelp) +
  geom_boxplot(aes(x=price, y=review_count),
               color="black",fill="white")

plotly::????(bxplot)
```

```{r}
a <- ggplot(data = yelp) +
  geom_boxplot(aes(x=price, y=review_count), 
               fill = "white", color = "black")

b <- ggplot(data = yelp) +
  geom_boxplot(aes(x=review_count, y=price), 
               fill="white", color="black") 

gridExtra::grid.arrange(a, b)
```

Of course, you can use `scale_fill_manual()` to use your custom color. To do this, however, you need to have specified `fill` inside `aes()`. If it is not specified at all or is specified outside of `aes()`, custom color won't work.

```{r}
yelp %>%
  st_join(census %>% st_transform(crs = st_crs(yelp))) %>% 
  separate(col = NAME, into=c("tract","county","state"),sep=", ") %>% 
  drop_na(county) %>% 
  ggplot() +
  geom_boxplot(aes(x = price, y=review_count, fill = price), 
               color = "black") +
  facet_wrap(~county) +
  scale_fill_brewer(palette = "Blues")
```

Violin plot is yet another plot for visualizing the distribution of variables. While a box plot allows you to see where the upper/lower fences are and where the median and quartiles are, a violin plot allows you to see the concentration of observations at a certain bins (or range if you will) of values.

```{r}
vplot <- ggplot(data = yelp %>%
                  st_join(census %>% st_transform(crs = st_crs(yelp))) %>% 
                  separate(col = NAME, into=c("tract","county","state"),sep=", ") %>% 
                  drop_na(county)) +
  geom_violin(aes(x = price, y=review_count, fill = price), 
               color = "black") +
  facet_wrap(~county) +
  scale_fill_brewer(palette = "Blues")

plotly::ggplotly(vplot)
```

# So, do restaurants in wealthy neighborhoods get higher Yelp ratings?

As we've seen so far, data visualization is an indispensable tool for urban analysts. If we have some time left in the class, think about socioeconomic/demographic variables in American Community Survey that may associate with review count, rating, and/or price of Yelp data. Then, get the data, plot them, and draw insights from them.

### Association between household income and Yelp ratings.
```{r warning=F}
census_yelp %>% 
  mutate(review_count_cut = cut(review_count, breaks = quantile(review_count, prob = c(0,0.5,0.75,1)), include.lowest=TRUE)) %>%
  ggplot(data = ., aes(x = ????, y = ????)) +
  geom_point(mapping = aes(color = review_count_cut)) +
  scale_color_manual(values = c("gray50", "orange", "red"), labels = c("0 - 50th", "50th- 75th", "75th - 100th")) +
  labs(x = "Annual Household Income", y = "Yelp Rating", color = "Review Count (discrete)", title = "Household Income vs. Rating") +
  ggdark::dark_theme_gray() +
  
  # ------------------------------------------------------------------
  # This line of code adds the correlation analysis result to the plot
  ggpubr::stat_cor(method = "pearson", label.x = 160000, label.y = 1.5)
```

### Association between White population(%) and Yelp ratings.

```{r warning=F}
census_yelp %>% 
  mutate(review_count_cut = cut(review_count, breaks = quantile(review_count, prob = c(0,0.5,0.75,1)), include.lowest=TRUE)) %>%
  mutate(pct_white = race.white / pop) %>% 
  ggplot(data = ., aes(x = ????, y = ????)) +
  geom_point(mapping = aes(color = review_count_cut)) +
  scale_color_manual(values = c("gray50", "orange", "red"), labels = c("0 - 50th", "50th- 75th", "75th - 100th")) +
  labs(x = "Proportion of White Residents", y = "Yelp Rating", color = "Review Count (discrete)", title = "Proportion of White Residents vs. Rating") +
  ggdark::dark_theme_gray() + 
  
  # ------------------------------------------------------------------
  # This line of code adds the correlation analysis result to the plot
  ggpubr::stat_cor(method = "pearson", label.x = 0.6, label.y = 1.5) 
```
