---
title: "Computer Vision"
author: "Originally written by Bon Woo Koo & Subhro Guhathakurta; modified by Uijeong Hwang"
date: '2024-10-30'
output:
  rmdformats::downcute:
    downcute_theme: "chaos"
---

<style type="text/css">
  body{
  font-family: Arial;
  }
</style>

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
library(osmdata)
library(sfnetworks)
library(units)
library(sf)
library(tidygraph)
library(tmap)
library(here)
ttm()
```

We are going to do quick analyses using the semantic segmentation model results.

## Merge the data processed from Colab back to R. 

If you are having trouble running the Colab script, you may use these files: [geojson](https://raw.githubusercontent.com/ujhwang/urban-analytics-2024/main/Lab/module_4/edges_azi_chamblee.geojson) and [csv](https://raw.githubusercontent.com/ujhwang/urban-analytics-2024/main/Lab/module_4/seg_output.csv).

```{r}
# Load the sampled points.
edges_azi <- st_read(????) # path to your "edges_azi_{city_name}.geojson" file

# Download the output from the computer vision models.
seg_output <- read.csv(????) # path to your "seg_output.csv" file

# Join them back to the `edges_azi` object that was used to download images.
edges_seg_output <- edges_azi %>% 
  inner_join(seg_output, by=c("node_id"="img_id"))
```

The numeric values in each object column represent the number of **pixels** in images predicted as that object. Check out the first few rows and see which objects comprise the majority of streetscapes.

```{r}
head(edges_seg_output)
```

## Calculate Greenness

Let's define greenness as **the sum of tree, grass, and plant**.
```{r}
edges_seg_output %<>% # `a %<>%` is the same as `a <- a %>%`; you need to import `magrittr` to use this pipe
  mutate(greenness = ????,
         pct_greenness = greenness/(640*640))

# Map!
t1 <- tm_basemap("OpenStreetMap")+
  tm_shape(edges_seg_output) + 
  tm_dots(col = "pct_greenness", style="quantile", palette = 'viridis')

t2 <- tm_basemap(leaflet::providers$Esri.WorldImagery) +
  tm_shape(edges_seg_output %>% st_bbox() %>% st_as_sfc()) + 
  tm_borders(lwd = 2, col = 'white') +

tmap_arrange(t1, t2, sync = T)
```
  
  
## Calculate Building-to-street Ratio

Let's calculate building-to-street ratio. Major 'building' objects include ***building*** and ***house***. Major 'street' objects include ***road***, ***sidewalk***, and ***car***.

```{r}
edges_seg_output %<>% 
  mutate(b2s_ratio = ????) %>% 
  mutate(b2s_ratio = case_when(
    b2s_ratio >= 1 ~ 1, # Let's set the upper limit as 1.
    TRUE ~ b2s_ratio))

t1 <- tm_basemap("OpenStreetMap")+
  tm_shape(edges_seg_output) + 
  tm_dots(col = "b2s_ratio", style="quantile", palette = 'viridis')

t2 <- tm_basemap(leaflet::providers$Esri.WorldImagery) +
  tm_shape(edges_seg_output %>% st_bbox() %>% st_as_sfc()) + 
  tm_borders(lwd = 2, col = 'white')

tmap_arrange(t1, t2, sync = T)
```
  
  
## Violin plots comparing the segmentation results by type of road

Are there significant differences between road types?
```{r}
edges_seg_output %<>%
  filter(!is.na(highway)) %>% 
  mutate(highway = factor(highway, levels = c('unclassified','residential','tertiary',
                                              'secondary','primary','trunk'))) %>%
  mutate(pct_building = building/(640*640),
         pct_sky = sky/(640*640),
         pct_road = road/(640*640),
         pct_sidewalk = sidewalk/(640*640))

edges_seg_output %>% 
  pivot_longer(cols = c(b2s_ratio, pct_greenness, pct_building, 
                        pct_sky, pct_road, pct_sidewalk), 
               names_to = 'variable', 
               values_to = "value") %>% 
  ggplot(mapping = aes(x = highway, y = value)) +
  geom_violin(fill = "#c7ffd6") +
  stat_summary(fun=mean, geom="point", size=1, color="red")+
  coord_flip() +
  theme_bw() +
  facet_wrap(~variable, scales = "free_x", nrow = 2)
```










