---
title: "Tidy Data"
author: "Originally written by Bon Woo Koo & Subhro Guhathakurta; modified by Uijeong Hwang"
institute: "Georgia Institute of Technology"
date: "2023/9/14"
output:
  xaringan::moon_reader:
    css: ["default", "assets/sydney-fonts.css", "assets/sydney.css"]
    seal: false
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '4:3' # alternatives '16:9' or '4:3' or others e.g. 13:9
---
```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#1c5253",
  header_font_google = google_font("Open Sans"),
  text_font_google   = google_font("Source Sans Pro", "400", "400i"),
  code_font_google   = google_font("Fira")
)

nice_table <- function(df, height="150px"){
  return(
    df %>% kable() %>% 
      kable_styling(latex_options="scale_down", font_size=12) %>% 
      scroll_box(width = "100%", height = height)
    )
}
```

```{r, include=F}
library(xaringanthemer)
library(sf)
library(tidyverse)
library(leaflet)
library(tmap)
library(kableExtra)
library(tidycensus)
tmap_mode('view')
```

class: middle, inverse

# Preparing your Data
## Data Wrangling

.font100[
Subhro Guhathakurta & Uijeong Hwang

9/14/2023
]
---
# Data are messy
When you read textbooks on statistics or online tutorials on Urban Analytics, they often use data that’s already nice and clean. They often do not show you all the hassle they went through to clean the data

--
### Unprocessed Data are messy and have all sorts of issues

--
For example:
--
<ol>
<li> Rows and columns do not represent observations and variables, respectively
<li> Duplicated or redundant rows/columns
<li> multiple variables in one column
<li> Missing data
<li> Unstructured data
</ol>
and the list goes on..
---
# Issue 1 - Rows and columns do not represent observations and variables

## The opposite of "messy" data is "tidy" data
--

.center[
In tidy data,

“Each variable must have its own column.”

“Each observation must have its own row.”

“Each value must have its own cell.”

(from Chapter 12.2 in R for Data Science)
]
--

- <font color=blue>You need to decide what is your "variable" and what is an individual "record"</font>
---
# Rearranging columns and rows
## Wide vs. Long forms
```{r, include=F, message=F}
census_api_key(Sys.getenv("census_api"))

suppressMessages(
census_wide <- get_acs(geography="tract", state="GA", variables=c(hhincome='B19019_001', race.tot = "B02001_001"), 
                       county="Fulton", year=2019, output="wide", geomtry=TRUE))
suppressMessages(
census_long <- get_acs(geography="tract", state="GA", variables=c(hhincome='B19019_001', race.tot = "B02001_001"), 
                       county="Fulton", year=2019, geomtry=TRUE))
```
.small[
.pull-left[

with `output="wide"`:

```{r}
census_wide %>% 
  head() %>%  
  nice_table("350px")
```
]

.pull-right[
with `output="long"`:

```{r}
census_long %>% 
  head() %>% 
  nice_table("350px")
```
]
]
---
## Wide vs. Long forms

.small[
```{r}
longer <- census_wide %>% 
  pivot_longer(cols = hhincomeE:race.totM, # Cols to be affected
               names_to = c("variable"), # Name for the label column
               values_to = c("value")) # Name for the value column
longer
```
]

---
## Wide vs. Long forms

.small[
```{r}
wider <- longer %>% 
  pivot_wider(id_cols = c(GEOID, NAME), 
              names_from = c("variable"),
              values_from = c("value")) 
wider
```
]

---
# Issue 2 - Duplicated or redundant rows/columns
<ol>
<li> Duplicated rows is a common issue when using API to acquire data
<li> Multiple functions exist in R to deal with duplicates 
<li> Most frequently used ones include:
<ol>
- duplicated() and distinct()


```{r}
dupl_df <- data.frame(name = c("A", "A", "B", "C", "C", "C", "D"),
                      GPA = c(3.5, 3.5, 4.0, 2.0, 3.0, 3.0, 2.0)) 
duplicated(dupl_df$name)
```
---
# !duplicated()

```{r}
# Duplicates in column "name" removed.
dupl_df[!duplicated(dupl_df$name),]
```
---
# distinct()
```{r}
# Returns a vector, not data frame
dupl_df %>% 
  distinct(name) # Try adding .keep_all = TRUE argument 
```
```{r}
# Returns a data frame
dupl_df %>% 
  distinct(name, GPA)
```
---
# Issue 3 - Multiple variables in one column
## This issue comes in two broad types.

1. Different variables can be concatenated into a long string

A very common example from Census data: you fill often find something like <b>Census Tract 9501, Appling County, Georgia</b> in a single column

It contains at least three variables: Tract, County, and State

We can break the string down into pieces using <font color=red>separate()</font> from <font color=blue>tidyr</font> package

---
# separate()
```{r}
# A character vector to split
onecol_df <- data.frame(labels = c('a1','b_2','c_3_2','d_4_1'))
# split the character at _
onecol_df %>% separate(col = "labels", sep = "_", into = c("alphabet", "numeric")) 
```
---
# The second type:
##One column in a data frame can contain another data frame or a list
This data structure is frequently found when we convert something like a JSON, which can have nested structure, into a data frame. 

In such cases you need to <font color=blue>flatten</font> the data frame using jsonlite::flatten() which will automatically generate columns that concatenate the name of the column with the nested columns inside

--

Even after flattening, there could be list-columns remaining. These will have to be addressed individually according to your needs 
---
# Issue 4 - Missing Values
In R, NA is used to represent missing data 

Many core functions in R does not work properly when NA is in the vector (e.g., try mean(c(NA, 1, 2, 3)))

```{r}
avg <- mean(c(NA, 10, 25, 3))
avg
```
So it is imperative that NAs are taken care of

is.na() is a Base R function that returns a logical object (TRUE/FALSE)

drop.na() is part of the tidyr package that drops columns with NAs

If drop.na() is used without specifying columns then only those rows that do not have NAs will be retained
---
.small[
* is.na()
```{r}
# This is the same toy_df from above
toy_df <- data.frame(name = c("John", "Jane", "Mary"), 
                     treatment_a = c(NA, 16, 3),
                     treatment_b = c(2, 11, 1),
                     treatment_c = c(6, 12, NA))

# Dropping NA using is.na()
toy_df %>% 
  filter(!is.na(treatment_a))
```
* drop.na()
```{r}
# This check across all columns and drops all rows that have at least one NA.
toy_df %>% 
  drop_na()
```
]

---
## Saving files

R has two native data formats:
* rds: stores a *single* R object
* RData: stores *multiple* R objects

---
## Saving files

### (1) Read/write .rds
* **`write_rds()`, `read_rds()`**: Save and load a single object.

```{r eval=F}
library(tidyverse)
a <- c(1,2,3)

a %>% write_rds(file = 'data/a.rds')
a <- read_rds('data/a.rds')

a %>% write_rds(file = here('data', 'a.rds'))
a <- read_rds(here('data', 'a.rds'))
```

---
## saving files

### (2) Read/write .RData
* **`save()`**: Save multiple objects.
* **`save.image()`**: Save all the objects in the environment.  
* **`load()`**: Load .RData file. Do not use **`<-`** (as .RData file stores the original object name as well).

```{r eval=F}
a <- c(1,2,3)
b <- c(3,4,5)

save(a, b, file = 'data.RData') # never use piping for 'save()'
load('data.RData')

save.image(file = 'data_all.RData') # saves everything
load('data_all.RData')
```

---
## Anonymous function

.small[
* When using **apply()** or **map()**, you can provide an existing or a custom-made function.
* Similar to lambda in Python, R has anonymous function. 
* Anonymous function is a function defined on the fly and disappears after execution.

.pull-left[
```{r}
map(1:5, # input vector
    function(x) x + 1) # anonymous function
```
]

.pull-right[
```{r}
map(1:5, # input vector
    function(x){ # anonymous function with {}
      out <- (x + 1)*x
      return(out)
    })
```
]
]

---
## Anonymous function

.small[
* map() and other variants has a nice syntax that make the code simple. 
* Instead of declaring `function(x)`, you can use a tilde (~) to indicate that it is anonymous function.
* `x`s inside the anonymous function needs to be preceded by a period (.). See the example below.

```{r}
map(1:5, # input
    ~(.x + 1)*.x ) # tilde replaces function().  # x is preceded by a period
```
]

---
## Existing data frame into a sf object
* You can convert a data frame with lng/lat into a sf object. This can be done using **st_as_sf()**. 
* The word 'as' indicates that it converts an *existing* object to sf rather than creating one from scratch.

.pull-left[
```{r eval=F}
# A data frame with XY info
point_df <- data.frame(
  x = c(-84.3991, -84.4010, -84.3899), 
  y = c(33.7770, 33.7748, 33.7777))

# st_as_sf
point_sf <- point_df %>% 
  st_as_sf(coords = c("x", "y"), 
           crs = 4326)

tm_shape(point_sf) + tm_dots()
```
]

.pull-right[
```{r, echo=F, out.width="100%"}
# A data frame with XY info
point_df <- data.frame(
  x = c(-84.3991, -84.4010, -84.3899), 
  y = c(33.7770, 33.7748, 33.7777))

# st_as_sf
point_sf <- point_df %>% 
  st_as_sf(coords = c("x", "y"), 
           crs = 4326)

tm_shape(point_sf) + tm_dots()
```
]


---
## Detecting string - str_detect/grepl

* Some may download two Yelp categories at once.
* Rows for category A and B will be mixed.
* To count how many As and Bs are there respectively, you need to search strings.

grepl(.red[pattern], .blue[string])
```{r}
a <- c("yoga studio", "health gym", "pizza", "YoGa")
grepl("yoga", a)
```

str_detect(.blue[string], .red[pattern])
```{r}
str_detect(a, "yoga")
```
